---
title: "Dataquest-Forest Fire"
author: "Aulia Larasati"
date: "5/27/2021"
output: pdf_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(RColorBrewer)
library(tidyverse)
library(rnaturalearth) 
library(rnaturalearthdata)
knitr::opts_chunk$set(echo = TRUE)
```

1. Reading data 
```{r}
forest_fire <- read.csv("forestfires (1).csv")
```
2. Data cleaning: 
Create variables that are easily remembered 
```{r}
#Data cleaning, renaming variable
forest_fire <- forest_fire %>% 
  rename(x_spatial = X, 
         y_spatial = Y, 
         fine_fuel_moist_code = FFMC, 
         duff_moist = DMC, 
         initial_spread_index = ISI, #spread of fire
         relative_humid = RH) 

```

3. Dataquest question:  
a. What does a single row represent?  
A single row represent the weather and climate surrounding a certain point in 
the map at a certain time.  

b. With what I know about fires, how might each of the variables relate to the
fires themselves? This might involve looking up each variable quickly on a search 
engine and getting a better grasp on what they are.  

The weather / meteorological condition (temperature, humidity, rain, and wind) 
are used to predict fires to inform firefighting management.   
FWI is the Canadian system to rate fire danger, it contains FFMC, DMC, DC, and ISI.
FFMC : Fine fuel moisture code = moisture content that ignition and fire spread
DMC = fire intensity, 
ISI = spread of fire 
RH = relative humidity. 

c. Convert day and month 
```{r}
forest_fire <- forest_fire %>% 
  mutate(month = factor(month, levels = c("jan", "feb", "mar", "apr", "may", 
                             "jun", "jul", "aug", "sep", "oct","nov", "dec")),
         day = factor(day, levels = c("mon", "tue", "wed", "thu", "fri", "sat", 
                                      "sun"))
         )
```

d. Create a tibble that counts the number of forest by month 
```{r}
forest_fire %>% 
  group_by(month) %>% 
  summarize(n_fire = n()) %>% 
  ggplot(aes(x = month, y = n_fire)) + geom_col(aes(fill = n_fire)) + 
  labs(x = "month", y = "number of forest fires", 
       title = "Forest Fires are highest in August and September") + 
  scale_fill_continuous(high = "darkred")
```
d. Create a tibble that counts the number of fire by the day of the week 
```{r}
forest_fire %>% 
  group_by(day) %>% 
  summarize(n_fire = n()) %>% 
  mutate(day = fct_reorder(day, n_fire)) %>% 
  ggplot(aes(x = day, y = n_fire)) + geom_col(aes(fill = day)) + 
  labs(x = "day", y = "number of forest fires", 
       title = "The Number of Forest Fires are more common on Friday - Saturday ") + 
  scale_fill_brewer(palette = "RdYlGn")

```

Forest fires are more common in the months of August and September.
Specifically in August the number of fires can reach 200. When we look 
at the data based on days, the number of forest fire is highest in Friday and 
Sunday

e. Relationship between data each column with month 
```{r}
forest_fire_long <- forest_fire %>% 
  pivot_longer(fine_fuel_moist_code:area, names_to = "variable", 
               values_to = "value")

#vizualization based on month
month_fire <- forest_fire_long %>% 
  ggplot(aes(x = month, y = value, group = month)) + 
  coord_flip() + 
  geom_point() + facet_wrap(~variable, scales = "free_x")
month_fire
```
The value for DC, duff moist, FFMC, and ISI is highest in September and August. 
The values mentioned are used to measure forest fire and its spread in Canada. 

```{r}
day_fire <- 
  forest_fire_long %>% 
  ggplot(aes(x = day, y = value, group = day)) + 
  coord_flip() + 
  geom_point() + facet_wrap(~variable, scales = "free_x")
day_fire
```
```{r}
area_forest <- forest_fire %>% 
   pivot_longer(fine_fuel_moist_code:rain, names_to = "variable", 
               values_to = "value") 

area_forest %>% 
  ggplot(aes(x = value, y = area, group = variable)) + 
  geom_point() +  facet_wrap(~variable, scales = "free")
```
Looking at the relationship between area and other related variables. the widest
scope of forest fire occur when there has been no rain, relatively low humidity, 
high FFMC, DC, temperature, and with wind strength around 2.5 to 5. However, 
it is hard to see the general patern because the datapoints are consentrated at
the bottom. 

```{r}
area_forest %>% 
  filter(area > 1  & area < 300) %>% 
  ggplot(aes(x = area, y = value, group = variable)) + 
  geom_point() +  facet_wrap(~variable, scales = "free")

summary(area_forest)
```



