---
title: "Dataquest-NY Public School 2011 Survey"
author: "Aulia Larasati"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

#Key information from data dictionary

- dbn = unique identifier for schools  
- sch_type = school type (elementary, middle, high)
- location = school name 
(_s : student, _t = teacher _p = parents, _tot = total)
- nr = number of eligible 
- saf = safety and respect score
- com = communication score 
- eng = engagement score 
- aca = academic expectation 

We can merge surveys to combined csv using "DBN" variable. 

# Read data 
1. Combined dataset: https://data.world/dataquest/nyc-schools-data/workspace/file?filename=combined.csv 
2. NYC School Survey: https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8 
```{r}
combined <- read_csv("combined.csv")
general_ed <- read_tsv("masterfile11_gened_final.txt")
district75 <- read_tsv("masterfile11_d75_final.txt")
```

# Data Cleaning 
```{r}
#rename DBN and filter to only high school in community school 
general_ed_HS <- general_ed %>% 
  rename(DBN = dbn) %>% 
  filter(schooltype == "High School") %>% 
  select(DBN, schoolname, rr_s:aca_tot_11)


district75_clean <- district75 %>% 
  rename(DBN = dbn) %>%
  select(DBN, schoolname, rr_s:aca_tot_11)

#combining district 75 with general education  
school_survey <- general_ed_HS %>% 
  bind_rows(district75_clean)

#school_survey with combined df
school_survey <- combined %>% 
  left_join(school_survey, by = "DBN")

#rename race for future aesthetics of plots 
school_survey <- school_survey %>% 
  rename(asian = asian_per, white = white_per, 
         black = black_per, hispanic = hispanic_per, 
         male = male_per, female = female_per)
```

# Correlation Analysis: School Quality, Demographic, and Academic Success
1. Relationship between demographic and school quality with avg sat score
```{r}
#creating correlation matrix
cor_matrix1 <- school_survey %>% 
  select(avg_sat_score, avg_class_size, asian:female, 
         saf_p_11:aca_tot_11) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  as.tibble(rownames = "variable")
```

```{r}
#variable with strong correlation (r-squared) above +/- 0.25 
cor_matrix1 %>% 
  select(variable, avg_sat_score) %>% 
  filter(avg_sat_score <= -0.25 | avg_sat_score >= 0.25)
```
 

## plot 
1. Relationship between average SAT score with demographic: a higher share of 
black and hispanic student negatively correlates with the average SAT score. 
While a higher share of asian and white students have a strong positive 
correlation with a higher SAT scores. 
```{r}
sat_race <- school_survey %>% 
  select(DBN, avg_sat_score, asian, white, black, hispanic) %>% 
  pivot_longer(asian:hispanic, names_to = "race", values_to = "percent")

sat_race %>% 
  filter(!is.na(avg_sat_score) | !is.na(percent)) %>%
  ggplot(aes(x = percent, y = avg_sat_score, group = race, color = race)) + 
  geom_point() + geom_smooth() +  facet_wrap(~race) + 
  labs(title = "Relationship between Race and SAT score in NY Highschool in 2011",
      x = "share of race in school (%)", y = "Average SAT Score" )
  
```

2. Relationship between students academic expectation with SAT scores: There 
is a positive correlation between students academic expectation with the SAT 
results in all subjects. 
```{r}
sat_aca <- school_survey %>% 
  rename(`Total Avg. SAT` = avg_sat_score) %>% 
  select(aca_s_11, `SAT Critical Reading Avg. Score`,`SAT Math Avg. Score`, 
        `SAT Writing Avg. Score`, `Total Avg. SAT`) %>% 
  pivot_longer(2:5, names_to = "SAT subject", values_to = "Score" )

sat_aca %>% 
  filter(!is.na(aca_s_11) | !is.na(Score)) %>% 
  ggplot(aes(x = aca_s_11, y = Score, 
             color = `SAT subject`)) + geom_point() + geom_smooth() + 
  facet_wrap(~`SAT subject`, scales = "free")
```

3. Relationship between class size and SAT scores: The majority of schools have 
a class size between 20 to 30 and the average class size is positively correlated
with a higher total average sat scores. 
```{r}
school_survey %>% 
  ggplot(aes(x = avg_class_size, y = avg_sat_score, color = avg_class_size)) + 
  geom_point() + geom_smooth() + xlim(20,40)
```

# Differences in student, parent, and teacher responses to survey 

## Data Cleaning
```{r}
#reshape
school_survey_long <- school_survey %>% 
  pivot_longer(saf_p_11:aca_tot_11, names_to = "survey_var", 
               values_to = "score")

#create new variable str_sub
school_survey_long <- school_survey_long %>% 
  mutate(respondent = str_sub(survey_var, 4,6), 
         variable = str_sub(survey_var, 1,3))

#change respondent from code to parents, students, teacher
school_survey_long <- school_survey_long %>% 
  mutate(respondent = case_when(respondent == "_p_" ~ "parent", 
                                respondent == "_t_" ~ "teacher", 
                                respondent == "_s_" ~ "student"))
```

## plot 
```{r}
school_survey_long %>% 
  filter(!is.na(respondent)) %>% 
  ggplot(aes(x = variable, y = score, fill = respondent)) + geom_boxplot()
```
Parents tend give a higher score than student in when asked on 
school's academic expectation, communication, engagement, and safety and respect.
While as the students who experience the school environment, they tend 
to provide a lower score compared to both parents and teachers. In particular, 
the difference is highest in communication and safety. 
